---
title: "R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---
***
## Fourier Series

The Fourier series for a periodic function of any period $(T)$ is given by:

$$
\begin{aligned}
f(t) = \frac {a_0}{2}+\sum_{i=1}^\infty\{a_n cos(\frac{2n\pi t}{T})+b_n sin(\frac{2n\pi t}{T})\}
\end{aligned}
$$ 

Where $a_n$ and $b_n$ are given by: 

$$
\begin{array}
{rrr}
a_n =  \frac {2}{T}  \sum_{\frac {-T}{2}}^\frac{T}{2}  f(t)cos(\frac{2n\pi t}{T}) & & & & & & for & n = 0, 1, 2 ...
\end{array}
$$

$$
\begin{array}
{rrr}
b_n =  \frac {2}{T}  \sum_{\frac {-T}{2}}^\frac{T}{2}  f(t)sin(\frac{2n\pi t}{T}) & & & & & & for & n = 1, 2, 3 ...
\end{array}
$$

*** 
#### Example 1

A periodic function is given by:   

$$
\begin{array}
{rrr}
f(t) = \{ & 0 & -2<t<0 & & &f(t+4)=f(t) \\
 \{ & t & 0<t<2 & &
\end{array}
$$  
  
* Period given is $T=4$   
* Limits for this period are $-2<t<2$   

This periodic function can be plotted as a piecewise periodic function.  

Using __*Ft.func*__ the periodic function can be coded using __*ifelse*__ commands.   
The __*Ft.func*__ can then be plotted using the __*plot.piecewise.fnc*__.  

__*plot.piecewise.fnc*__ inputs:  
  
* __*FUN = ... *__  Name given to the peridoic function eg: __*Ft.func*__    
* __*.piece.range = ... *__ Limits of the period eg: __*c(-2, 2)*__    
* __*.pieces = ... *__ Number of pieces to be included in the piecewise plot eg: __*c(-2,-1,0,1)*__ (4 pieces)  
* __*.resolution = ...*__ Number of points plotted eg: __*101*__

```{r}
library(tidyverse)
source("scr_functions.R")
Ft.func = function(t){
  ifelse(t<0, 0, t)
}

plot.piecewise.fnc(.FUN = Ft.func, .piece.range = c(-2, 2), .pieces = c(-2,-1,0,1), .resolution = 101)

```

*** 
#### Example 2

A periodic function is given by: 

$$
\begin{array}
{rrr}
f(t) = \{ & t^2 & -3<t<3 & & &f(t+6)=f(t) \\
\end{array}
$$  
  
* Period given is $T=6$  
* Limits for this period are $-3<t<3$

```{r}
Ft.func2 = function(t){
  t^2
}

plot.piecewise.fnc(.FUN = Ft.func2, .piece.range = c(-3, 3))
```
*** 
#### Example 3

A periodic function is given by: 

$$
\begin{array}
{rrr}
f(t) = \{ & -t & -2<t<0 & & &f(t+4)=f(t) \\
 \{ & 7 & 0<t<2 & &
\end{array}
$$
* Period given is $T=4$  
* Limits for this period are $-2<t<2$

```{r}
Ft.func3 = function(t){
  ifelse(t<0, -t, 7)
}

plot.piecewise.fnc(.FUN = Ft.func3, .piece.range = c(-2, 2))
```
*** 
#### Example 4

A periodic function is given by: 

$$
\begin{array}
{rrr}
f(t) = \{ & -t-3 & -5<t<-3 & & &f(t+13)=f(t) \\
\{ & t+3 & -3<t<0 & & \\
\{ & -2t+3 & 0<t<3 & &\\
\{ & t-6 & 3<t<8 & &
\end{array}
$$  
* Period given is $T=13$  
* Limits for this period are $-5<t<8$  

Periodic functions with mulitple definitions for different parts of the domain require more __*ifelse*__ commands.

```{r}
Ft.func4 = function(t){
  ifelse(t<=-3, -t-3,
         ifelse(t>-3 & t<=0, t+3,
                ifelse(t>0 & t<=3, -2*t+3,
                       t-6)))
}

plot.piecewise.fnc(.FUN = Ft.func4, .piece.range = c(-5, 8), .pieces = seq(-2,2))
```

*** 
#### Example 5: Functions of period $2\pi$

The Fourier series for a periodic function of period  $T=2\pi$ has the general form:

$$
\begin{aligned}
f(t) = \frac {a_0}{2}+\sum_{i=1}^\infty\{a_n cos(nt)+ b_n sin(nt)\}
\end{aligned}
$$ 

A piecewise plot for the function $sin(t)$ can be seen: 

```{r}
Ft.func5 = function(t){
  sin(t)
}

plot.piecewise.fnc(.FUN = Ft.func5, .piece.range = c(0, 2*pi), .pieces = 0)
```


*** 
#### Example 6

A periodic function (__*example 1*__) is given by:  

$$
\begin{array}
{rrr}
f(t) = \{ & 0 & -2<t<0 & & &f(t+4)=f(t) \\
 \{ & t & 0<t<2 & &
\end{array}
$$  
This periodic function is plotted as a piecewise periodic function: 

```{r echo=FALSE}

plot.piecewise.fnc(.FUN = Ft.func, .piece.range = c(-2, 2), .pieces = c(-2,-1,0,1), .resolution = 101)

```

The Fourier Series $f(t)$ including constants $a_0$, $a_n$ and $b_n$ are calculated.  

$$
\begin{array}
{rrr}
a_0 = & 1 & &  \\
&&& \\
a_n = & \{ & \frac{-4}{\pi^2n^2} & n=1,3,4... \\
& \{ & 0 & n=2,4,6... 
\end{array}
$$

$$
\begin{array}
{rrr}
b_n = &\frac{-2}{\pi n} cos(\pi n)&= \frac{-2}{\pi n}(-1)^n&
\end{array}
$$

The constants are then coded using using __*ifelse*__ commands.   
The Fourier series can then be plotted using the __*plot.fourier.series*__ function.    

```{r}
a0 = 1
an.fn = function(n) {
  ifelse(as.logical(n%%2), -4/(n^2*pi^2), 0)
}

bn.fn = function(n) {
  2/(n*pi) * ifelse(as.logical(n%%2), 1, -1)
}

plot.fourier.series(.T = 4, .a0 = a0, .an.fun = an.fn, .bn.fun = bn.fn,
                    .t.range = c(-10,6), .n.max = 10, .resolution = 300)

```

Increasing __*.n.max*__ (number of terms in Fourier Series) gives a closer approximation of $f(t)$

__*.n.max = 20*__
```{r echo=FALSE}

plot.fourier.series(.T = 4, .a0 = a0, .an.fun = an.fn, .bn.fun = bn.fn,
                    .t.range = c(-10,6), .n.max = 20, .resolution = 300)

```

__*.n.max = 30*__
```{r echo=FALSE}

plot.fourier.series(.T = 4, .a0 = a0, .an.fun = an.fn, .bn.fun = bn.fn,
                    .t.range = c(-10,6), .n.max = 30, .resolution = 300)

```
